---
title: "CoSIBS Project: LLM Model Comparison for Agentic AI Benchmarking"
author: "Zengtao Xing"
date: "July 26, 2026"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    code_folding: hide
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This report summarizes results the agentic AI benchmarking pipeline where the underlying LLM was varied while all other
pipeline parameters were held constant. The LLM is
responsible for retrieving and configuring candidate predictive models for a
hyperlipidemia classification task, which are then evaluated for performance.

**Area under the receiver operating characteristic curve (AUROC)** is used
as the primary performance metric rather than raw accuracy, because AUROC
accounts for both the true-positive and false-positive rate and is more
robust to the small number of hyperlipidemia diagnoses.

# Methods

## Experimental Design

A one-factor-at-a-time design was used, varying only the LLM across four
levels: **GPT-5.2**, **GPT-5.4**, **GPT-5.4-mini**, and **GPT-5.6-sol**. All
other pipeline parameters (temperature, prompt template, and number of
candidate models retrieved) were held fixed across runs. At least ten
independent runs were collected per LLM.

## Statistical Analysis Plan

Within each run, the LLM retrieves and fits three candidate models
(e.g., logistic regression, random forest, XGBoost). Because these models are benchamrked together, they are not independent
observations. To have independence, metrics were averaged **within each run** to satisfy the independence assumption of one-way ANOVA.

The following statistical tests were used to find differences in LLM means/effects:

1. One-way ANOVA
2. Welch's ANOVA
3. Kruskal-Wallis test
4. Games-Howell pairwise differences

# Packages

```{r packages}
# install.packages(c("dplyr", "car", "ggplot2", "rstatix", "knitr"))
library(dplyr)
library(car)
library(rstatix)
library(ggplot2)
library(knitr)
library(stringr)
```

# Data Compilation

## Building the Run-Level Dataframe

The raw JSON results from each run were compiled into a single long-format
dataframe (with the help of Claude), with one row per run &times; algorithm combination. Missing
metrics (e.g., runs where XGBoost failed to execute) are encoded as
`NA_real_`.

```{r build-dataframe}
df <- tribble(
  ~group, ~run, ~algorithm, ~accuracy, ~auroc, ~brier, ~f1, ~precision, ~recall, ~threshold,

  # ---------------- gpt-5.2 (11 runs) ----------------
  "gpt-5.2", 1, "logistic_regression", 0.34552845528455284, 0.5355780022446689, 0.36821516241928626, 0.4755700325732899, 0.3230088495575221, 0.9012345679012346, 1.83541123016041e-09,
  "gpt-5.2", 1, "random_forest",       0.4146341463414634,  0.5789375233819678, 0.217147662601626,   0.4782608695652174, 0.3384615384615385, 0.8148148148148148, 0.28,
  "gpt-5.2", 1, "xgboost",             0.4146341463414634,  0.6134680134680135, 0.24187880674759857, 0.5,                0.34782608695652173,0.8888888888888888, 0.0330415703356266,

  "gpt-5.2", 2, "lightgbm",       0.32926829268292684, 0.5, 0.22085258062076668, 0.4954128440366972, 0.32926829268292684, 1.0, 0.32789115646258504,
  "gpt-5.2", 2, "random_forest",  0.32926829268292684, 0.5, 0.22085309905425582, 0.4954128440366972, 0.32926829268292684, 1.0, 0.3277142857142856,
  "gpt-5.2", 2, "xgboost",        0.32926829268292684, 0.5, 0.22085258058078594, 0.4954128440366972, 0.32926829268292684, 1.0, 0.32789117097854614,

  "gpt-5.2", 3, "logistic_regression",              0.32926829268292684, 0.5, 0.22085258497426277, 0.4954128440366972, 0.32926829268292684, 1.0, 0.3278895767348184,
  "gpt-5.2", 3, "random_forest_classifier",         0.32926829268292684, 0.5, 0.22085093453156943, 0.4954128440366972, 0.32926829268292684, 1.0, 0.32976870748299314,
  "gpt-5.2", 3, "xgboost_extreme_gradient_boosting",0.32926829268292684, 0.5, 0.2208659009171721,  0.4954128440366972, 0.32926829268292684, 1.0, 0.32536742091178894,

  "gpt-5.2", 4, "logistic_regression", 0.33739837398373984, 0.45536849981294425, 0.32117345417474885, 0.48253968253968255, 0.3247863247863248, 0.9382716049382716, 0.0003545654918688194,
  "gpt-5.2", 4, "random_forest",       0.33739837398373984, 0.5012345679012346,  0.23162142873931058, 0.4953560371517028, 0.3305785123966942, 0.9876543209876543, 0.12,
  "gpt-5.2", 4, "xgboost",             0.42276422764227645, 0.49614665170220723, 0.2815598044474337,  0.4365079365079365, 0.3216374269005848, 0.6790123456790124, 0.08168642967939377,

  "gpt-5.2", 5, "gradient_boosting_machines_gbm", 0.3983739837398374,  0.5780396558174336, 0.21918745310570673, 0.5066666666666667,  0.3470319634703196, 0.9382716049382716, 0.16076323382918842,
  "gpt-5.2", 5, "random_forest",                  0.3333333333333333,  0.5552936775158998, 0.2467989268292683,  0.49693251533742333, 0.3306122448979592, 1.0,                0.002,
  "gpt-5.2", 5, "support_vector_machines_svm",    0.3333333333333333,  0.421623643845866,  0.220891956168234,   0.47435897435897434, 0.3203463203463203, 0.9135802469135802, 0.32809249986954586,

  "gpt-5.2", 6, "logistic_regression",              0.32926829268292684, 0.5, 0.22085085928269002, 0.4954128440366973, 0.32926829268292684, 1.0, 0.32884976419786244,
  "gpt-5.2", 6, "random_forest",                    0.32926829268292684, 0.5, 0.22085292471876558, 0.4954128440366973, 0.32926829268292684, 1.0, 0.32777142857142844,
  "gpt-5.2", 6, "xgboost_extreme_gradient_boosting",0.32926829268292684, 0.5, 0.2208659009171721,  0.4954128440366973, 0.32926829268292684, 1.0, 0.32536742091178894,

  "gpt-5.2", 7, "lightgbm",      0.34146341463414637, 0.535465768799102,  0.23900812525134155, 0.49375,             0.3305439330543933, 0.9753086419753086, 0.03437740890524628,
  "gpt-5.2", 7, "random_forest", 0.3333333333333333,  0.5331088664421998, 0.2717037733926648,  0.47770700636942676, 0.3218884120171674, 0.9259259259259259, 0.007630683217167498,
  "gpt-5.2", 7, "xgboost",       0.34146341463414637, 0.5307519640852975, 0.24667952690105371, 0.4873417721518987,  0.3276595744680851, 0.9506172839506173, 0.03524316847324371,

  "gpt-5.2", 8, "logistic_regression", 0.4024390243902439, 0.4919191919191919, 0.22996542312446155, 0.43243243243243246, 0.3146067415730337, 0.691358024691358,   0.32793834325133847,
  "gpt-5.2", 8, "random_forest",       0.5,                0.4853722409277964, 0.2643795822742859,  0.3626943005181347,  0.3125,             0.43209876543209874, 0.32433177550566156,
  "gpt-5.2", 8, "xgboost",             NA_real_,           NA_real_,           NA_real_,             NA_real_,            NA_real_,           NA_real_,            NA_real_,

  "gpt-5.2", 9, "logistic_regression", 0.32926829268292684, 0.5, 0.22085258497426277, 0.4954128440366973, 0.32926829268292684, 1.0, 0.3278895767348184,
  "gpt-5.2", 9, "random_forest",       0.32926829268292684, 0.5, 0.22085708027958334, 0.4954128440366973, 0.32926829268292684, 1.0, 0.3267392290249434,
  "gpt-5.2", 9, "xgboost",             NA_real_,            NA_real_, NA_real_,            NA_real_,            NA_real_,            NA_real_, NA_real_,

  "gpt-5.2", 10, "logistic_regression", 0.33739837398373984, 0.5236812570145903, 0.39786863063220207, 0.4792332268370607,  0.3232758620689655,  0.9259259259259259, 3.1432218136504514e-10,
  "gpt-5.2", 10, "random_forest",       0.4268292682926829,  0.6154133931911709, 0.21306691656064658, 0.5017667844522968,  0.35148514851485146, 0.8765432098765432, 0.2927942718273594,
  "gpt-5.2", 10, "xgboost",             0.33739837398373984, 0.5236812570145903, 0.39786863063220207, 0.4792332268370607,  0.3232758620689655,  0.9259259259259259, 3.1432218136504514e-10,

  "gpt-5.2", 11, "logistic_regression", 0.3333333333333333,  0.559670781893004,  0.2253631159775527,  0.49382716049382713, 0.3292181069958848,  0.9876543209876543, 0.0143180644288395,
  "gpt-5.2", 11, "random_forest",       0.32926829268292684, 0.4832772166105499, 0.2418378739749688,  0.4891640866873065,  0.32644628099173556, 0.9753086419753086, 0.018126118326118323,

  # ---------------- gpt-5.4 (10 runs) ----------------
  "gpt-5.4", 1, "logistic_regression",       0.573170731707317,   0.5737373737373737, 0.2214851401798883,  0.378698224852071,  0.36363636363636365, 0.3950617283950617, 0.338357061612245,
  "gpt-5.4", 1, "random_forest_classifier",  0.6544715447154471,  0.6118967452300785, 0.21413978760789093, 0.33070866141732286,0.45652173913043476, 0.25925925925925924,0.37796831443919376,
  "gpt-5.4", 1, "support_vector_machine_svm",0.6016260162601627,  0.5509913954358399, 0.21958582247046737, 0.3,                0.3559322033898305,  0.25925925925925924,0.33046439975722747,

  "gpt-5.4", 2, "logistic_regression",       0.35365853658536583, 0.5378226711560045, 0.3722690696099948,  0.4752475247524752, 0.32432432432432434, 0.8888888888888888, 1.2296560393423142e-07,
  "gpt-5.4", 2, "random_forest_classifier",  0.34146341463414637, 0.5832023943135054, 0.21767134146341463, 0.5,                0.3333333333333333,  1.0,                0.16,
  "gpt-5.4", 2, "support_vector_machine_svm",0.45528455284552843, 0.5013842124953236, 0.2208005078052702,  0.45081967213114754,0.3374233128834356,  0.6790123456790124, 0.3225869330277011,

  "gpt-5.4", 3, "logistic_regression", 0.33739837398373984, 0.5439955106621773, 0.3732893736176403,  0.4792332268370607, 0.3232758620689655, 0.9259259259259259, 5.366680802513263e-10,
  "gpt-5.4", 3, "random_forest",       0.3943089430894309,  0.5989898989898991, 0.2146861788617886,  0.4879725085910653, 0.3380952380952381, 0.8765432098765432, 0.25,

  "gpt-5.4", 4, "logistic_regression", 0.3333333333333333,  0.514702581369248,  0.3788509426484114,  0.4875,               0.3263598326359833, 0.9629629629629629, 1.269859791000437e-07,
  "gpt-5.4", 4, "random_forest",       0.4065040650406504,  0.611821922933034,  0.21248026196928638, 0.5067567567567568,   0.3488372093023256, 0.9259259259259259, 0.24,
  "gpt-5.4", 4, "xgboost",             NA_real_,            NA_real_,           NA_real_,            NA_real_,             NA_real_,           NA_real_,           NA_real_,

  "gpt-5.4", 5, "logistic_regression",                       0.3902439024390244,  0.5357276468387578, 0.22053332736607434, 0.506578947368421,   0.3452914798206278,  0.9506172839506173, 0.3156685435290104,
  "gpt-5.4", 5, "random_forest_classifier",                  0.3333333333333333,  0.5675645342312009, 0.21968364689134648, 0.49382716049382713, 0.3292181069958848,  0.9876543209876543, 0.304241801238824,
  "gpt-5.4", 5, "support_vector_machine_svm_with_rbf_kernel",0.32926829268292684, 0.4264870931537598, 0.22087622659977688, 0.4891640866873065,  0.32644628099173556, 0.9753086419753086, 0.3261981992225341,

  "gpt-5.4", 6, "logistic_regression",              0.34552845528455284, 0.4804339693228582, 0.29830434594819943, 0.4888888888888889,  0.32905982905982906, 0.9506172839506173, 0.00427476128215051,
  "gpt-5.4", 6, "random_forest",                    0.4065040650406504,  0.6183314627759072, 0.21249932249322492, 0.5034013605442177,  0.3474178403755869,  0.9135802469135802, 0.25,
  "gpt-5.4", 6, "xgboost_extreme_gradient_boosting", NA_real_,           NA_real_,            NA_real_,            NA_real_,            NA_real_,            NA_real_,           NA_real_,

  "gpt-5.4", 7, "logistic_regression", 0.32926829268292684, 0.5282454171343061, 0.3696306939856307,  0.48598130841121495, 0.325,               0.9629629629629629, 1.263803073190228e-11,
  "gpt-5.4", 7, "random_forest",       0.44308943089430897, 0.5954732510288065, 0.21541554878048783, 0.4981684981684982,  0.3541666666666667,  0.8395061728395061, 0.27,
  "gpt-5.4", 7, "xgboost",             NA_real_,            NA_real_,           NA_real_,            NA_real_,            NA_real_,            NA_real_,           NA_real_,

  "gpt-5.4", 8, "random_forest", 0.4024390243902439, 0.629105873550318, 0.219960298102981, 0.49829351535836175, 0.3443396226415094, 0.9012345679012346, 0.3333333333333333,

  "gpt-5.4", 9, "gradient_boosting_machine_gbm", 0.42276422764227645, 0.6133931911709689, 0.22362765881556687, 0.5202702702702703, 0.3581395348837209,  0.9506172839506173, 0.14140628878976916,
  "gpt-5.4", 9, "random_forest",                 0.32926829268292684, 0.5634118967452302, 0.22985075475525646, 0.4954128440366973, 0.32926829268292684, 1.0,                0.2830188679245283,
  "gpt-5.4", 9, "support_vector_machine_svm",    0.32926829268292684, 0.4427609427609427, 0.2341625309011664,  0.4954128440366973, 0.32926829268292684, 1.0,                0.26168224299065423,

  "gpt-5.4", 10, "logistic_regression", 0.32926829268292684, 0.5401795735129068, 0.22068278694994947, 0.4954128440366973, 0.32926829268292684, 1.0,                0.0,
  "gpt-5.4", 10, "random_forest",       0.32926829268292684, 0.5057987280209502, 0.2207713031685464,  0.4827586206896552, 0.3235294117647059,  0.9506172839506173, 0.31903847483737335,
  "gpt-5.4", 10, "xgboost",             NA_real_,            NA_real_,           NA_real_,            NA_real_,            NA_real_,            NA_real_,           NA_real_,

  # ---------------- gpt-5.4-mini (12 runs) ----------------
  "gpt-5.4-mini", 1, "logistic_regression",              0.3821138211382114, 0.5274971941638609, 0.38986949818620353, 0.4864864864864865, 0.33488372093023255, 0.8888888888888888, 1.424986430444194e-07,
  "gpt-5.4-mini", 1, "random_forest",                    0.3617886178861789, 0.5929292929292929, 0.21504075203252032, 0.4852459016393443, 0.33035714285714285, 0.9135802469135802, 0.23,
  "gpt-5.4-mini", 1, "xgboost_extreme_gradient_boosting", 0.3699186991869919, 0.6024691358024691, 0.24045158922672272, 0.5016077170418006, 0.3391304347826087, 0.9629629629629629, 0.022018438205122948,

  "gpt-5.4-mini", 2, "logistic_regression",      1.0,                 1.0, 4.7003810652642314e-05, 1.0,               1.0, 1.0,                0.966575527074772,
  "gpt-5.4-mini", 2, "random_forest_classifier", 0.967479674796748,   1.0, 0.010997865853658535,    0.948051948051948,1.0, 0.9012345679012346, 0.79,
  "gpt-5.4-mini", 2, "xgboost",                  1.0,                 1.0, 7.820214705134276e-06,   1.0,               1.0, 1.0,                0.9960005879402161,

  "gpt-5.4-mini", 3, "random_forest_classifier",         0.3617886178861789, 0.582865693976805, 0.2166183830171635,  0.47840531561461797, 0.32727272727272727, 0.8888888888888888, 0.23,
  "gpt-5.4-mini", 3, "support_vector_machine_svm",       0.3333333333333333, 0.5170594837261504, 0.220709922689941,  0.4906832298136646,  0.3278008298755187,  0.9753086419753086, 0.31669521856777666,
  "gpt-5.4-mini", 3, "xgboost_extreme_gradient_boosting", 0.524390243902439,  0.621623643845866, 0.23767435550689697, 0.5224489795918368,  0.3902439024390244,  0.7901234567901234, 0.10514029860496521,

  "gpt-5.4-mini", 4, "logistic_regression",       0.3821138211382114, 0.5323606434717546, 0.35839209098213287, 0.4864864864864865,  0.33488372093023255, 0.8888888888888888, 9.408835347140976e-07,
  "gpt-5.4-mini", 4, "random_forest_classifier",  0.33739837398373984,0.6319117096894875, 0.21068028455284551, 0.48580441640378547, 0.326271186440678,   0.9506172839506173, 0.185,
  "gpt-5.4-mini", 4, "support_vector_machine_svc",0.3333333333333333, 0.4963711185933408, 0.22108348725898697, 0.49693251533742333, 0.3306122448979592,  1.0,                0.3124379867944233,

  "gpt-5.4-mini", 5, "logistic_regression",      0.3333333333333333,  0.518742985409652,  0.23429039056100148, 0.47770700636942676, 0.3218884120171674, 0.9259259259259259, 0.15227189667355107,
  "gpt-5.4-mini", 5, "random_forest_classifier", 0.32926829268292684, 0.6430976430976432, 0.21121187895212282, 0.4954128440366973,  0.32926829268292684,1.0,                0.0,
  "gpt-5.4-mini", 5, "xgboost",                  0.3902439024390244,  0.6159371492704826, 0.25114116072654724, 0.5098039215686274,  0.3466666666666667, 0.9629629629629629, 0.018346205353736877,

  "gpt-5.4-mini", 6, "logistic_regression", 1.0, 1.0, 0.00033987544786001975, 1.0, 1.0, 1.0, 0.9885811608520114,
  "gpt-5.4-mini", 6, "random_forest",       1.0, 1.0, 0.022320189701897017,   1.0, 1.0, 1.0, 0.6233333333333333,
  "gpt-5.4-mini", 6, "xgboost",             1.0, 1.0, 1.4807117905177016e-05, 1.0, 1.0, 1.0, 0.9920083284378052,

  "gpt-5.4-mini", 7, "logistic_regression",     0.328, 0.47590011614401856, 0.26848607018475795, 0.4939759036144578, 0.328, 1.0,                2.4488805072302775e-05,
  "gpt-5.4-mini", 7, "random_forest_classifier",0.32,  0.5215229384436701,  0.22699364554358703,  0.4817073170731707, 0.32113821138211385,0.9634146341463414, 0.155,
  "gpt-5.4-mini", 7, "support_vector_machine_svm",0.328,0.49085365853658536,0.2206616954121636,   0.4939759036144578, 0.328, 1.0,                0.0,

  "gpt-5.4-mini", 8, "lightgbm",             0.3617886178861789,  0.5468013468013467, 0.27767913734004696, 0.49517684887459806, 0.3347826086956522,  0.9506172839506173, 0.006651109552742605,
  "gpt-5.4-mini", 8, "logistic_regression",  0.34959349593495936, 0.4719790497568275, 0.3974243583558975,  0.4936708860759494,  0.33191489361702126, 0.9629629629629629, 3.95141905266612e-10,
  "gpt-5.4-mini", 8, "random_forest",        0.32926829268292684, 0.5901608679386456, 0.21764308943089428, 0.4954128440366973,  0.32926829268292684, 1.0,                0.15666666666666668,

  "gpt-5.4-mini", 9, "lightgbm_light_gradient_boosting_machine", 0.3902439024390244,  0.5967078189300411, 0.2723242294004079,  0.5033112582781457,  0.3438914027149321, 0.9382716049382716, 0.003819564009478319,
  "gpt-5.4-mini", 9, "random_forest",                            0.3983739837398374,  0.5881032547699213, 0.21570727190605238, 0.5,                 0.34418604651162793,0.9135802469135802, 0.25333333333333335,
  "gpt-5.4-mini", 9, "xgboost_extreme_gradient_boosting",        0.4186991869918699,  0.6104002992891883, 0.23111432790756226, 0.49110320284697506, 0.345,              0.8518518518518519, 0.0876733735203743,

  "gpt-5.4-mini", 10, "gradient_boosting_classifier",  0.46747967479674796, 0.5842873176206509, 0.22229298417534318, 0.45643153526970953, 0.34375,             0.6790123456790124, 0.21828088505299315,
  "gpt-5.4-mini", 10, "random_forest_classifier",      0.4065040650406504,  0.5829405162738496, 0.21403333333333335, 0.4930555555555556,  0.34299516908212563,0.8765432098765432, 0.25,
  "gpt-5.4-mini", 10, "support_vector_machine_svm",    0.34552845528455284, 0.49270482603815935,0.22109188797652304, 0.47896440129449835, 0.32456140350877194,0.9135802469135802, 0.31923949276373764,

  "gpt-5.4-mini", 11, "logistic_regression", 0.6707317073170732, 0.39169472502805835, 0.3043714353536765,  0.0,                  0.0,                0.0,                  1.0,
  "gpt-5.4-mini", 11, "random_forest",       0.6747967479674797, 0.4132809577254022,  0.2588581197689133,  0.024390243902439025, 1.0,                0.012345679012345678, 0.8793333333333332,
  "gpt-5.4-mini", 11, "xgboost",             0.6585365853658537, 0.5349046015712682,  0.2310888022184372,  0.045454545454545456, 0.2857142857142857, 0.024691358024691357, 0.5407688021659851,

  "gpt-5.4-mini", 12, "lightgbm",            0.32926829268292684, 0.5,                0.22094107130159188, 0.4954128440366973, 0.32926829268292684, 1.0, 0.0,
  "gpt-5.4-mini", 12, "logistic_regression", 0.34146341463414637, 0.509090909090909, 0.22089868121324369, 0.5,                0.3333333333333333,  1.0, 0.3378637861773547,
  "gpt-5.4-mini", 12, "xgboost",             0.32926829268292684, 0.5,                0.2209661900997162,  0.4954128440366973, 0.32926829268292684, 1.0, 0.0,

  # ---------------- gpt-5.6-sol (14 runs) ----------------
  "gpt-5.6-sol", 1, "gradient_boosting_classifier",  0.36585365853658536, 0.6128694350916574, 0.22525513057194058, 0.5031847133757962, 0.33905579399141633, 0.9753086419753086, 0.10037818229745232,
  "gpt-5.6-sol", 1, "random_forest_classifier",      0.4186991869918699,  0.6025065469509914, 0.2197634247685411,  0.5051903114186851, 0.35096153846153844, 0.9012345679012346, 0.31861207184612944,
  "gpt-5.6-sol", 1, "support_vector_machine_svm",    0.34959349593495936, 0.5448559670781894, 0.2208542679200526,  0.5031055900621118, 0.3360995850622407,  1.0,                0.32637586057303336,

  "gpt-5.6-sol", 2, "random_forest",             0.4349593495934959,  0.6004489337822672, 0.21811419049987366, 0.508833922261484,   0.3564356435643564, 0.8888888888888888, 0.3063960695993427,
  "gpt-5.6-sol", 2, "support_vector_machines_svm",0.3333333333333333,  0.4298540965207632, 0.2208872868366391,  0.49693251533742333, 0.3306122448979592,  1.0,                0.32548737989039495,
  "gpt-5.6-sol", 2, "xgboost",                    0.3333333333333333,  0.6039655817433596, 0.2158094201672745,  0.49693251533742333, 0.3306122448979592,  1.0,                0.24208248253823123,

  "gpt-5.6-sol", 3, "lightgbm",                  0.34959349593495936, 0.5881780770669659, 0.2177907638698843,  0.5,                 0.33472803347280333, 0.9876543209876543, 0.28559501825195294,
  "gpt-5.6-sol", 3, "random_forest_classifier",  0.5040650406504065,  0.5877291432846988, 0.21725778600166692, 0.47413793103448276, 0.36423841059602646, 0.6790123456790124, 0.3157288833760119,
  "gpt-5.6-sol", 3, "xgboost",                   0.3821138211382114,  0.588327721661055,  0.21598365928530833, 0.5,                 0.34080717488789236, 0.9382716049382716, 0.2568023750788275,

  "gpt-5.6-sol", 4, "lightgbm",      0.4146341463414634,  0.5976056864945753, 0.2157117532980871,  0.5102040816326531,  0.352112676056338,   0.9259259259259259, 0.29646886221710395,
  "gpt-5.6-sol", 4, "random_forest", 0.43089430894308944, 0.5932285821174709, 0.21969385883132875, 0.48905109489051096, 0.3471502590673575,  0.8271604938271605, 0.3194669320300703,
  "gpt-5.6-sol", 4, "xgboost",       0.3780487804878049,  0.607482229704452,  0.2175384029795331,  0.5080385852090032,  0.34347826086956523, 0.9753086419753086, 0.30635464984681693,

  "gpt-5.6-sol", 5, "logistic_regression",       0.32926829268292684, 0.5802469135802469, 0.21804532752403682, 0.4954128440366973, 0.32926829268292684, 1.0,                0.2762326199869898,
  "gpt-5.6-sol", 5, "random_forest_classifier",  0.32926829268292684, 0.5859708193041525, 0.21902640551000935, 0.4891640866873065, 0.32644628099173556, 0.9753086419753086, 0.2856239506779278,
  "gpt-5.6-sol", 5, "xgboost",                   0.45934959349593496, 0.6042648709315376, 0.21378032405784134, 0.5055762081784386, 0.3617021276595745,  0.8395061728395061, 0.28380755817037995,

  "gpt-5.6-sol", 6, "logistic_regression",              0.32926829268292684, 0.5500187055742611, 0.21866090881257597, 0.49230769230769234, 0.32786885245901637, 0.9876543209876543, 0.3033822086805061,
  "gpt-5.6-sol", 6, "random_forest",                    0.3902439024390244,  0.5951365506921062, 0.21964855273950887, 0.49324324324324326, 0.3395348837209302,  0.9012345679012346, 0.3167681451027182,
  "gpt-5.6-sol", 6, "xgboost_extreme_gradient_boosting", 0.44715447154471544, 0.638757949869061,  0.23377574262917178, 0.5142857142857142,  0.36180904522613067, 0.8888888888888888, 0.06463021039962769,

  "gpt-5.6-sol", 7, "random_forest",              0.3333333333333333,  0.5814814814814815, 0.21833813190906373, 0.49693251533742333, 0.3306122448979592,  1.0,                0.2706751285437187,
  "gpt-5.6-sol", 7, "support_vector_machines_svm",0.34146341463414637, 0.542536475869809,  0.22077543910979924, 0.5,                 0.3333333333333333,  1.0,                0.32386220501629037,
  "gpt-5.6-sol", 7, "xgboost",                    0.43902439024390244, 0.5950617283950618, 0.21528082497087467, 0.49635036496350365, 0.35233160621761656, 0.8395061728395061, 0.29463663758410064,

  "gpt-5.6-sol", 8, "gradient_boosting_classifier",  0.34552845528455284, 0.6182566404788626, 0.22348518070881737, 0.4984423676012461, 0.3333333333333333,  0.9876543209876543, 0.06015786755892765,
  "gpt-5.6-sol", 8, "random_forest_classifier",      0.4024390243902439,  0.562102506546951,  0.21896060779873247, 0.4693140794223827, 0.33163265306122447, 0.8024691358024691, 0.30381435843353516,
  "gpt-5.6-sol", 8, "support_vector_machine_svc",    0.3333333333333333,  0.4249158249158249, 0.22120587510778292, 0.4906832298136646, 0.3278008298755187,  0.9753086419753086, 0.32022127148107793,

  "gpt-5.6-sol", 9, "lightgbm",             0.3699186991869919,  0.5994014216236437, 0.21742766841196803, 0.5016077170418006,  0.3391304347826087,  0.9629629629629629, 0.28847862911830524,
  "gpt-5.6-sol", 9, "logistic_regression",  0.32926829268292684, 0.5354283576505798, 0.22105620644173568, 0.4954128440366973,  0.32926829268292684, 1.0,                0.25071557962646995,
  "gpt-5.6-sol", 9, "xgboost",              0.3333333333333333,  0.6004489337822672, 0.21653570225886148, 0.49693251533742333, 0.3306122448979592,  1.0,                0.24760659085861514,

  "gpt-5.6-sol", 10, "lightgbm",      0.3821138211382114,  0.5790497568275346, 0.2157447774774505,  0.5032679738562091,  0.3422222222222222,  0.9506172839506173, 0.2992059303670567,
  "gpt-5.6-sol", 10, "random_forest", 0.4268292682926829,  0.583426861204639,  0.21926015320495712, 0.49097472924187724, 0.3469387755102041,  0.8395061728395061, 0.3158919237572334,
  "gpt-5.6-sol", 10, "xgboost",       0.43089430894308944, 0.6065843621399177, 0.21646310377287453, 0.5035460992907801,  0.35323383084577115, 0.8765432098765432, 0.2980906116650275,

  "gpt-5.6-sol", 11, "logistic_regression", 0.4349593495934959,  0.5634493078937524, 0.2188427822263489,  0.505338078291815,   0.355,               0.8765432098765432, 0.3024836444784287,
  "gpt-5.6-sol", 11, "random_forest",       0.44715447154471544, 0.6014216236438459, 0.2165208658599439,  0.5,                 0.35602094240837695, 0.8395061728395061, 0.3011396575466156,
  "gpt-5.6-sol", 11, "xgboost",             0.34146341463414637, 0.6212495323606435, 0.25430671116041487, 0.4968944099378882,  0.33195020746887965, 0.9876543209876543, 0.002605562098324299,

  "gpt-5.6-sol", 12, "logistic_regression",     0.3333333333333333,  0.5473999251777029, 0.22011091089612547, 0.4875,              0.3263598326359833, 0.9629629629629629, 0.3055809985918014,
  "gpt-5.6-sol", 12, "random_forest_classifier",0.43089430894308944, 0.5722783389450056, 0.21839916245382496, 0.4927536231884058,  0.3487179487179487, 0.8395061728395061, 0.3024566820202103,
  "gpt-5.6-sol", 12, "xgboost",                 0.42276422764227645, 0.6268612046389824, 0.2381842781343656,  0.5170068027210885,  0.3568075117370892, 0.9382716049382716, 0.023647857829928398,

  "gpt-5.6-sol", 13, "lightgbm",      0.34552845528455284, 0.584287317620651,  0.21662766371383196, 0.4984423676012461,  0.3333333333333333,  0.9876543209876543, 0.2586239469137225,
  "gpt-5.6-sol", 13, "random_forest", 0.32926829268292684, 0.5847362514029181, 0.2179248092456341,  0.49230769230769234, 0.32786885245901637, 0.9876543209876543, 0.2499358553564878,
  "gpt-5.6-sol", 13, "xgboost",       0.3902439024390244,  0.571043771043771,  0.21827589302630465, 0.506578947368421,   0.3452914798206278,  0.9506172839506173, 0.28524874236446,

  "gpt-5.6-sol", 14, "lightgbm",      0.4146341463414634,  0.5889263000374112, 0.21739755900161709, 0.5167785234899329,  0.3548387096774194,  0.9506172839506173, 0.2976807679711833,
  "gpt-5.6-sol", 14, "random_forest", 0.32926829268292684, 0.5802095024317246, 0.21804067812066452, 0.4954128440366973,  0.32926829268292684, 1.0,                0.25890258802992694,
  "gpt-5.6-sol", 14, "xgboost",       0.3983739837398374,  0.6209502431724654, 0.2162652376859936,  0.5033557046979866,  0.3456221198156682,  0.9259259259259259, 0.289592519151415
)

df$group <- factor(df$group, levels = c("gpt-5.2", "gpt-5.4", "gpt-5.4-mini", "gpt-5.6-sol"))
```

## Data Overview

```{r counts-table}
run_counts <- df %>%
  group_by(group) %>%
  summarise(n_runs = n_distinct(run), n_rows = n())

kable(run_counts, caption = "Number of runs and total algorithm-level rows per model group.")

write.csv(df, "benchmark_results_long.csv", row.names = FALSE)
```

```{r}
# Raw frequency count
raw_freq <- df %>%
  count(group, algorithm, name = "n") %>%
  group_by(group) %>%
  slice_max(n, n = 1, with_ties = TRUE) %>%
  arrange(group, desc(n))

kable(raw_freq, caption = "Most frequent algorithm label (raw, unnormalized) per model group.")

# Grouping models with similar names to account for variations in LLM output
df <- df %>%
  mutate(
    model_family = case_when(
      str_detect(algorithm, "xgboost")                     ~ "xgboost",
      str_detect(algorithm, "lightgbm")                     ~ "lightgbm",
      str_detect(algorithm, "random_forest")                ~ "random_forest",
      str_detect(algorithm, "logistic")                     ~ "logistic_regression",
      str_detect(algorithm, "support_vector|svm|svc")        ~ "svm",
      str_detect(algorithm, "gradient_boosting")             ~ "gradient_boosting", # non-xgb/lgbm GBMs
      TRUE ~ algorithm  # fallback: leave anything unmatched as-is (flags new labels)
    )
  )

family_freq <- df %>%
  count(group, model_family, name = "n") %>%
  group_by(group) %>%
  mutate(pct = n / sum(n)) %>%
  arrange(group, desc(n))

# most frequent family per group (with ties kept)
top_family <- family_freq %>%
  group_by(group) %>%
  slice_max(n, n = 1, with_ties = TRUE)

kable(top_family, digits = 2,
      caption = "Most frequently chosen model family per model group (LLM-name variants collapsed).")

# full breakdown, useful for a stacked bar chart
kable(family_freq, digits = 2,
      caption = "Full model-family selection frequency by group.")
```


## Data By Each Run

To satisfy the independence assumption required for one-way ANOVA, metrics
are averaged across the two-to-three candidate algorithms fit within each
run, producing one independent observation per run.

```{r run-level-summary}
df_run <- df %>%
  group_by(group, run) %>%
  summarise(
    accuracy = mean(accuracy, na.rm = TRUE),
    auroc    = mean(auroc, na.rm = TRUE),
    brier    = mean(brier, na.rm = TRUE),
    f1       = mean(f1, na.rm = TRUE),
    .groups  = "drop"
  )
```

# Consistency and Uncertainty Metrics

To quantify uncertainty for each run, we examine the standard deviation (SD), coefficient of variation (CV = SD / mean), and interquartile range (IQR) of the mean AUROC.

```{r consistency-metrics}
consistency_summary <- df_run %>%
  group_by(group) %>%
  summarise(
    n          = n(),
    mean_auroc = mean(auroc, na.rm = TRUE),
    sd_auroc   = sd(auroc, na.rm = TRUE),
    cv_auroc   = sd(auroc, na.rm = TRUE) / mean(auroc, na.rm = TRUE),
    iqr_auroc  = IQR(auroc, na.rm = TRUE),
    min_auroc  = min(auroc, na.rm = TRUE),
    max_auroc  = max(auroc, na.rm = TRUE)
  )

kable(consistency_summary, digits = 3,
      caption = "Run-level AUROC summary statistics by model group.")
```

As a complementary notion of consistency, within-run agreement
across the candidate algorithms is also examined. For each run, the SD of
accuracy across its benchmarked models is computed, then
averaged across runs within a group. This captures whether the algorithms performed similarly.

```{r within-run-spread}
within_run_spread <- df %>%
  group_by(group, run) %>%
  summarise(sd_within_run_acc = sd(accuracy, na.rm = TRUE), .groups = "drop") %>%
  group_by(group) %>%
  summarise(mean_within_run_sd = mean(sd_within_run_acc, na.rm = TRUE))

kable(within_run_spread, digits = 3,
      caption = "Mean within-run SD of accuracy across candidate algorithms, by model group.")
```

**Observations:**

- GPT-5.4-mini shows a markedly higher SD and CV of AUROC than the other
  three model groups.
- GPT-5.4-mini's maximum observed AUROC is 1.0, and the corresponding runs
  also show Brier scores near zero.

# Hypothesis Testing

## Assumption Checks

Levene's test evaluates whether the assumption of equal variances across
groups holds. A significant result (p < 0.05) indicates that variances
differ meaningfully across groups, which would make a standard one-way
ANOVA inappropriate.

```{r levene-test}
levene_result <- leveneTest(auroc ~ group, data = df_run)
levene_result
```

The test returns a significant result, indicating that AUROC variance is
not homogeneous across the four model groups. Welch's ANOVA and the
Kruskal-Wallis test are therefore used alongside the standard ANOVA below.

## ANOVA

```{r anova-accuracy}
anova_fit_accuracy <- aov(accuracy ~ group, data = df_run)
summary(anova_fit_accuracy)
```

The standard one-way ANOVA on **accuracy** is not statistically
significant.

```{r anova-auroc}
anova_fit_auroc <- aov(auroc ~ group, data = df_run)
summary(anova_fit_auroc)
```

The standard one-way ANOVA on **AUROC** is not statistically
significant.

```{r welch-anova}
welch_result <- oneway.test(auroc ~ group, data = df_run, var.equal = FALSE)
welch_result
```

```{r kruskal-wallis}
kruskal_result <- kruskal.test(auroc ~ group, data = df_run)
kruskal_result
```

Based on Welch's ANOVA and the Kruskal-Wallis test for **AUROC**, we conclude that
not all group means are equal.

## Post-Hoc Comparisons

Games-Howell pairwise comparisons were used due to unequal variances and sample sizes.

```{r games-howell}
post_hoc <- games_howell_test(auroc ~ group, data = df_run)
kable(post_hoc, digits = 4,
      caption = "Games-Howell pairwise comparisons of mean AUROC by model group.")
```

**Interpretation:** GPT-5.4 returns significantly higher AUROC than GPT-5.2
(estimated difference of 0.04), and GPT-5.6-sol returns significantly
higher AUROC than GPT-5.2 (estimated difference of 0.06). No other pairwise
comparisons are significant, though the wide confidence
intervals involving GPT-5.4-mini likely reflect its elevated within-group
variance rather than a genuine absence of difference.

# Visualization

```{r plot-auroc, fig.width=7, fig.height=4.5}
group_ci <- df_run %>%
  group_by(group) %>%
  summarise(
    mean_auroc = mean(auroc, na.rm = TRUE),
    sd_auroc   = sd(auroc, na.rm = TRUE),
    n          = sum(!is.na(auroc)),
    se_auroc   = sd_auroc / sqrt(n),
    t_crit     = qt(0.975, df = n - 1),
    ci_low     = mean_auroc - t_crit * se_auroc,
    ci_high    = mean_auroc + t_crit * se_auroc,
    .groups    = "drop"
  )

ggplot(group_ci, aes(x = group, y = mean_auroc, color = group)) +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0.15, linewidth = 0.8) +
  geom_point(size = 3) +
  labs(
    title = "Mean AUROC by LLM (95% CI)",
    x = NULL, y = "Mean AUROC"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```


```{r plot-accuracy, fig.width=7, fig.height=4.5}
ggplot(df_run, aes(x = group, y = accuracy, color = group)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.6) +
  labs(title = "Run-Level Accuracy by Model Group",
       x = NULL, y = "Accuracy") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r plot-cv, fig.width=7, fig.height=4.5}
ggplot(consistency_summary, aes(x = group, y = cv_auroc, fill = group)) +
  geom_col() +
  labs(title = "Coefficient of Variation for AUROC by Model Group",
       x = NULL, y = "CV of AUROC") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r plot-proportion_models, fig.width=7, fig.height=4.5}
library(scales)

ggplot(family_freq, aes(x = group, y = pct, fill = model_family)) +
  geom_col(width = 0.65, color = "white", linewidth = 0.3) +
  geom_text(
    data = subset(family_freq, pct >= 0.08),
    aes(label = percent(pct, accuracy = 1)),
    position = position_stack(vjust = 0.5),
    color = "white", size = 3.2, fontface = "bold",
    show.legend = FALSE
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    panel.grid.major.x = element_blank()
  ) +
  labs(
    title = "Candidate Model Selection Frequency by LLM",
    x = NULL, y = "Proportion of Selections"
  )
```


# Conclusion

This experiment compared four LLMs (GPT-5.2, GPT-5.4, GPT-5.4-mini,
and GPT-5.6-sol) used to configure candidate models for agentic AI benchmarking, holding all other pipeline parameters
constant. From the statistical tests, we conclude that model means are **not** all
equal for mean AUROC. GPT-5.4 and GPT-5.6-sol each achieve significantly higher
mean AUROC than GPT-5.2, indicating a general trend of improved
benchmarking capability in more advanced model versions.

Beyond the mean comparisons, qualitative review of the run logs found that
GPT-5.2 had the highest rate of unsuccessful or degenerate runs (roughly
one in four runs), while GPT-5.6-sol had the highest rate of successful
runs (roughly seven in ten). Taken together with the AUROC results, this
suggests that more advanced LLMs are both more capable and more reliable
when used to configure predictive models in this pipeline.

Lastly, GPT-5.4-mini produced abnormally high variability across runs. This could be due to data leakage, but GPT-5.4-mini is worth further investigation.

## Limitations

- **Missing XGBoost results.** Several runs have `NA` values for XGBoost because the package was
  unsuccessfuly implemented in the execution environment for those runs, not because the
  model underperformed. These `NA`s were excluded from run-level averages, but they are a pipeline/environment issue rather than
  a genuine model-comparison signal and should be corrected at the source
  in future runs.
- **Sample size.** Group sizes (10-14 runs) are modest for detecting
  small-to-moderate effect sizes. Larger sample sizes (> 30 runs) could result in more significant pairwise comparisons.
